\documentclass[ebook,12pt,oneside,openany]{memoir}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}
\usepackage{url}

\begin{document}

\title{Why Greatness Cannot Be Planned}
\maketitle

We'll discover a source of optimism that embraces the uncertainty of the far future rather than fearing or denying it.
This journey begins by considering both the power and limitations of searching for novelty.
A deeper point is that novelty search will not always find what you want.
The myth of objective still shows its true colors when searching for an objective still performs worse than novelty search.

This strange paradox, where trying is a curse and not trying is a blessing, sets the stage for a more realistic understanding of what is achievable and how.
It means that ambitious goals can't reliably be achieved by trying-unless they are one stepping stone away, where they come within reach.
Otherwise, the only choice that remains is not trying. And while this treasure-hunting approach will not ensure reaching any particular objective,
what it will do is accumulate stepping stones that lead to unfamiliar places. The treasure hunter is an opportunistic explorer- searching for anything and everything of value, 
without a care for what might be found.

To be a treasure hunter, you have to collect as many stepping stones as you can, because you never know which one might lead somewhere valuable.
This insight may seem sad, that we're left with no sure compass, that all our efforts to create certainty and to search with purpose may be futile.
Perhaps search isn't about objectives but about something bigger. In that case, abandoning the false compasss can be liberating, opening up a new frontier.
If discovery without explicit objectives is the guiding light of natural evolution, of human innovation, then we might harness it for our own purposes.
Instead of something to fear it can be something to embrace.

We can actually build systems based on the non-objective principle. 
The possibility of building treasure-hunting systems becomes intriguing.
The way to unleash the treasure hunter is actually through separating people from each other, where people only interact by taking off from where someone else left.



\end{document}